---
title: "FSB Student Success Dashboard"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Marti Walstad and Alaina Edwards"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
#runtime: shiny
---
<style>                     
.navbar {
  background-color:#545454;
  border-color:black;
}
.navbar-brand {
color:white!important;
}
</style>   

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)

#package intialization
if(require(pacman)==FALSE) install.packages("pacman")
pacman::p_load(tidyverse, #for data manipulation
               plotly, #for interactive plots
               flexdashboard, #to make a dashboard
               fontawesome, #for icons
               DT #for better tables
               )

#read in data
data=readRDS("cleaned_FSB_Student_Success")
```



# Student Success Overview {data-orientation=columns}

## row {data-height=600}

### Data Exploration Overview

Data exploration includes creating this dashboard to answer Kirk Bogard's questions about student placement information. Specifically, the question we seek to answer is: 

**What majors have the best / worst placement results in terms of salary levels?**

We aim to provide statistics to be used in marketing that will be appealing to future prospective students about FSB majors as well as inform FSB rankings and attract and maintain relationship with companies recruiting from FSB. Our goal is to support overall FSB Brand Management. 

### Written Insights

Through this dashboard, we have graphically represented average salary outcome by major with a bar chart, and provided a count of total reponses by major to provide context. 

We also see to divide this data by exploring these outcomes with other variables as well. These factors provide more detail into salary outcomes and student major choices. We have created bar charts of average salary by major, organized by: 

**1) First Generation Student Status**

**2) Honor Student Status**

**3) Declared Minor**

**4) Year**

## row {data-height=600}

### Average Salary by Major
```{r, echo = FALSE}
sorted_data <- data %>%
  group_by(MajorAbbr) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create the bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary)) +
  geom_bar(stat = 'identity', position = 'dodge', color = '#C3142D', fill = '#C3142D') +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw()

# Adjust the y-axis to show more values
p <- p +
  scale_y_continuous(breaks = seq(0, max(sorted_data$AverageSalary), by = 10000), labels = scales::dollar_format(prefix = "$"))

# Convert to interactive plotly plot
ggplotly(p)
```
### Number of Responses For Salary by Major

```{r,echo=FALSE}
p1 = ggplot(data, aes(reorder(MajorAbbr,MajorAbbr, function(x)-length(x))))+
  geom_bar(color="#C3142D", fill = "#C3142D")+
  xlab("Major")+
  ylab("Number of Responses")+
  theme_bw()
ggplotly(p1)
```


# Salary Results with Factors {data-orientation=columns}
## row {data-height=600}

### First Gen Status, Salary by Major


```{r,echo=FALSE}
sorted_data <- data %>%
  group_by(MajorAbbr, FirstGen) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create a variable for the grouping of "FirstGen" and "Not FirstGen"
sorted_data$Group <- factor(sorted_data$FirstGen, levels = c(1, 0), labels = c("First Gen", "Not First Gen"))

# Create the clustered bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary, fill = Group)) +
  geom_bar(stat = 'identity', position = 'dodge', width = 0.5) +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw() +
  scale_fill_manual(values = c('First Gen' = "#C3142D", 'Not First Gen' = "#3A3B3C"))

# Adjust the y-axis to show more values
max_salary <- max(sorted_data$AverageSalary, na.rm = TRUE)  # Handle NA values
if (is.finite(max_salary)) {
  p <- p +
    scale_y_continuous(breaks = seq(0, max_salary, by = 10000), labels = scales::dollar_format(prefix = "$"))
}

# Convert to interactive plotly plot
ggplotly(p)  
```   
 
### Honor Student Status, Salary by Major

```{r,echo=FALSE}
sorted_data <- data %>%
  group_by(MajorAbbr, Honors) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create a variable for the grouping of "FirstGen" and "Not FirstGen"
sorted_data$Group <- factor(sorted_data$Honors, levels = c(1, 0), labels = c("Honors", "Not Honors"))

# Create the clustered bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary, fill = Group)) +
  geom_bar(stat = 'identity', position = 'dodge', width = 0.5) +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw() +
  scale_fill_manual(values = c('Honors' = "#C3142D", 'Not Honors' = "#3A3B3C"))

# Adjust the y-axis to show more values
max_salary <- max(sorted_data$AverageSalary, na.rm = TRUE)  # Handle NA values
if (is.finite(max_salary)) {
  p <- p +
    scale_y_continuous(breaks = seq(0, max_salary, by = 10000), labels = scales::dollar_format(prefix = "$"))
}

# Convert to interactive plotly plot
ggplotly(p) 
```


## row {data-height=600}

### Salary by Declared Minor


```{r,echo=FALSE}
sorted_data <- data %>%
  group_by(Minor) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create the bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(Minor, -AverageSalary), y = AverageSalary)) +
  geom_bar(stat = 'identity', position = 'dodge', color = '#C3142D', fill = '#C3142D', width = 0.5) +
  xlab("Minor") +
  ylab("Average Salary") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels by 45 degrees


# Adjust the y-axis to show more values
p <- p +
  scale_y_continuous(breaks = seq(0, max(sorted_data$AverageSalary), by = 10000), labels = scales::dollar_format(prefix = "$"))

# Convert to interactive plotly plot
ggplotly(p)
```

# Salary by Year {data-orientation=columns}
## Column {data-width=400}
### 2019
```{r,echo=FALSE}
filtered_data <- data[data$Year == 2019, ]


sorted_data <- filtered_data %>%
  group_by(MajorAbbr) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create the bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary)) +
  geom_bar(stat = 'identity', position = 'dodge', color = '#C3142D', fill = '#C3142D') +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw()

# Adjust the y-axis to show more values
p <- p +
  scale_y_continuous(breaks = seq(0, 80000, by = 10000), labels = scales::dollar_format(prefix = "$"))

# Convert to interactive plotly plot
ggplotly(p)
```
## Column {data-width=400}

### 2020
```{r,echo=FALSE}
filtered_data <- data[data$Year == 2020, ]


sorted_data <- filtered_data %>%
  group_by(MajorAbbr) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create the bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary)) +
  geom_bar(stat = 'identity', position = 'dodge', color = '#C3142D', fill = '#C3142D') +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw()

# Adjust the y-axis to show more values
p <- p +
  scale_y_continuous(breaks = seq(0, 80000, by = 10000), labels = scales::dollar_format(prefix = "$"))

# Convert to interactive plotly plot
ggplotly(p)
```
## Column {data-width=400}

### 2021
```{r,echo=FALSE}
filtered_data <- data[data$Year == 2021, ]


sorted_data <- filtered_data %>%
  group_by(MajorAbbr) %>%
  summarize(AverageSalary = mean(Salary)) %>%
  arrange(desc(AverageSalary))

# Create the bar plot with sorted MajorAbbr
p <- ggplot(sorted_data, aes(x = reorder(MajorAbbr, -AverageSalary), y = AverageSalary)) +
  geom_bar(stat = 'identity', position = 'dodge', color = '#C3142D', fill = '#C3142D') +
  xlab("Major") +
  ylab("Average Salary") +
  theme_bw()

# Adjust the y-axis to show more values
p <- p +
  scale_y_continuous(breaks = seq(0,80000, by = 10000), labels = scales::dollar_format(prefix = "$"))

# Convert to interactive plotly plot
ggplotly(p)
```

